---
author: "Daniel E. Duplisea"
output: md_document
---
## Carbon footprint for meetings and courses

Meetings, courses, workshops etc, means that participant travel, restaurant meals and stays in hotels cause carbon emissions over and above the carbon emissions of a participant not travelling. This carbon footprint calculator provides a rough but consistent means of determining carbon emmision incurred by a meeting and specifically in this case for TESA activities. The group Technical Expertise in Stock Assessment (TESA) organises courses and workshops each year for DFO employees to continually build the expertise in the Department of Fisheries and Oceans in stock assessment. It is essential that face-to-face meetings occur for many but not all of these activities.

In an effort to quantify TESA carbon emissions, to track them by year, activity and location, this carbon footprint calculator has been developed. It accounts only for travel related carbon emissions (transport, hotel, meals).

Carbon emissions (tonnes CO2) per km flying, driving, train or bus and emissions for a night in a hotel or average meal in a restaurant are taken from https://calculator.carbonfootprint.com/.

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

## Install and load the library
```{r install, eval=FALSE}
devtools::install_github("duplisea/TESAcarbon")
library(TESAcarbon)
```

```{r loadlibrary, include=FALSE}
library(TESAcarbon)
```

## Calculate the carbon footprint for an individual traveller

Five nights in a hotel, a 600 km plane trip (1200 km total return), 40 km of driving to and from airport solo and 15 meals

```{r personal}
C.f(hotel.nights=5,
    plane.distance=600*2,
    bustrain.distance= 0,
    car.distance=40,
    number.car.sharing=1,
    meals=15)
```


## Carbon emissions for a series of courses or meetings held by ICES

A mock example of a course offered by ICES is provided with the dataset ICES. It contains only the essential columns. You can have as many columns as you want but you need to make sure you have these ones and they have these exact names (case sensitive)

```{r ICES}
names(ICES)
```

You will likely want to create your own .csv file with the names above and all rows filled in (1 row per participant).
You would then import this and run the carbon footprint function on it. Here is the result of running the carbon.footprint.f function on the ICES data
```{r ICES.C}
carbon.footprint.f(ICES, "ICES training carbon footprint, mock example", list.out=F)
```

This creates a global map with all origin cities of all participants located on the map and the host cities for the courses/meetings are circled red. A bar graph of the carbon emissions per activity are shown with per capita values on each bar and the title contains the total emissions for all activities combined.

## Carbon emissions for TESA 2019-20 activities
This calculation is based on actual data though we are not sure of the final participation of activities yet to come, these are probably close

```{r TESA.C}
carbon.footprint.f(TESA19.20, "TESA carbon footprint 2019-20", list.out=F)
```

## A shiny implementation

You can run this as a shiny application on your webbrowser. You will need to install the package TESAcarbon (which requires devtools to install), shiny and shinybusy from CRAN (the code below checks to see if you have them and installs them if not) and then use the runGitHub command

```{r shinyapp, eval=FALSE}
if (!require('shiny')) install.packages('shiny'); library('shiny')
if (!require('shinybusy')) install.packages('shinybusy')
if (!require('devtools')) install.packages('devtools')
if (!require('TESAcarbon')) devtools::install_github('duplisea/TESAcarbon')
runGitHub(repo="TESAcarbon", username="duplisea", subdir="shiny")
```

## References

https://calculator.carbonfootprint.com/
