---
author: "Daniel E. Duplisea"
output: md_document
---
## Carbon footprint for TESA activities

This calculates and incremental carbon footprint for participants in TESA activities. From this you can calculate the total incremental carbon emissions from an activity (tonnes CO2).

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

## Install and load the library
```{r install, eval=FALSE}
devtools::install_github("duplisea/TESAcarbon")
library(TESAcarbon)
```

```{r loadlibrary, include=FALSE}
library(TESAcarbon)
```

## Calculate the incremental carbon footprint

Five nights in a hotel, a 1200 km return plane trip, 40 km of driving to and from airport solo and 15 meals

```{r personal}
C.f(hotel.nights=5,
    plane.distance=1200,
    bustrain.distance= 0,
    car.distance=40,
    number.car.sharing=1,
    meals=15)
```

Here is an example for a course being held in Ottawa with 27 participants from all regions. Most participants fly but two participants from Burlington decide to drive and share a car. Most participants stay five days but a couple stay only four. It is assumed that everyone except people from Ottawa have 14 meals. The Ottawa people have 1 meal which is the group supper. The course instructor is from Copenhagen and they have one more night's stay and 3 more meals.

```{r group}
#TESA.course= read.csv("TESA.course.csv")
TESA.course$C= C.f(hotel.nights=TESA.course$hotel.nights,
    plane.distance=TESA.course$air.distance*2,
    bustrain.distance= TESA.course$bustrain.distance*2,
    car.distance= TESA.course$car.distance*2,
    number.car.sharing=TESA.course$car.sharing,
    meals=TESA.course$meals)
TESA.course[,c(2,10)]
sum(TESA.course$C)
```

You will see that the Copenhagen trip is the most carbon heavy while the locals in Ottawa have the lowest footprint. Driving a large car from Burlington with two people is not very different than flying from Halifax or Moncton. The total incremental carbon footprint of TESA activity is 17.2 tonnes of C02.

## References

https://calculator.carbonfootprint.com/
