---
author: "Daniel E. Duplisea"
output: md_document
---
## Carbon footprint for TESA activities

This calculates and incremental carbon footprint for participants in TESA activities. From this you can calculate the total incremental carbon emissions from an activity (tonnes CO2).

Carbon emissions per km flying, driving, train or bus and emissions for a night in a hotel or average meal in a restaurant are taken from https://calculator.carbonfootprint.com/. Flying distances are generally taken as direct air distances between cities (https://canada-map.com/distance/). For smaller airports, where doubling back usually occurs then this is considered. For example Mont-Joli to St. John's requires that you fly southwest to Montreal before flying over Mont-Joli again northeast to get to St. John's. There we no effort to account for example that essentially straight flights would still have layovers at intermediate airports that could add km to the actual flying distance. The multiple possible routings can easily violate assumptions made in that kind of consideration.

Given assumptions, the most useful aspect of this is to keep calculations consistent and then look at changes from year to year in the carbon emissions.

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```

## Install and load the library
```{r install, eval=FALSE}
devtools::install_github("duplisea/TESAcarbon")
library(TESAcarbon)
```

```{r loadlibrary, include=FALSE}
library(TESAcarbon)
```

## Calculate the incremental carbon footprint

Five nights in a hotel, a 1200 km return plane trip, 40 km of driving to and from airport solo and 15 meals

```{r personal}
C.f(hotel.nights=5,
    plane.distance=1200,
    bustrain.distance= 0,
    car.distance=40,
    number.car.sharing=1,
    meals=15)
```

Here is an example for a course being held in Ottawa with 27 participants from all regions. Most participants fly but two participants from Burlington decide to drive and share a car. Most participants stay five days but a couple stay only four. It is assumed that everyone except people from Ottawa have 14 meals. The Ottawa people have 1 meal which is the group supper. The course instructor is from Copenhagen and they have one more night's stay and 3 more meals.

```{r group}
#TESA.course= read.csv("TESA.course.csv")
TESA.course$C= C.f(hotel.nights=TESA.course$hotel.nights,
    plane.distance=TESA.course$plane.distance*2,
    bustrain.distance= TESA.course$bustrain.distance*2,
    car.distance= TESA.course$car.distance*2,
    number.car.sharing=TESA.course$car.sharing,
    meals=TESA.course$meals)
TESA.course[,c(2,10)]
sum(TESA.course$C)
```

You will see that the Copenhagen trip is the most carbon heavy while the locals in Ottawa have the lowest footprint. Driving a large car from Burlington with two people is not very different than flying from Halifax or Moncton. The total incremental carbon footprint of TESA activity is 17.2 tonnes of C02.

## References

https://calculator.carbonfootprint.com/
